\documentclass[../main.tex]{subfiles}

\begin{document}
Urban air mobility (UAM) refers to air transport systems designed to move people and goods by air within and around dense city areas. It's crucuial to develop methods that ensure safety and maximize efficiency.
In \cite{pelegrin-2023} a mathematical formulation for reoptimization is presented.
The proposed algorithm is divided in three layers. In the first one the strategic deconflict is calculated prior to departure.
The second layer focuses on tactical deconfliction, with the goal of maintain separation between ongoing traffic.
As third and last layer is dedicated to collission avoidance, utilizing onboard technologies to evade imminent collision.
The paper concentrate on the second layer, which aims to avoid loss of separation due to random variations, priority traffic or non-cooperative intruder.
Additionaly three urban topologies are proposed  for testing, that are the Grid, Airport and Metroplex, that will be described  in detail in \ref{sec:maps}.\\
In \cite{portoleau-2024} a two-stage robust algorithm for the scheduling problem of UAM flight is presented.
The algorithm is based on Adversarial Bender Decomposition Method, and allows for solving the subproblems using heuristic.
One of the proposed heuristic involves fixing the order and solving the approximated subproblem with either a quadratic problem or with a linear problem with increasing piecewise-linear objective function.
On chapter \ref{chap:ExactAlgo} we will see another way to solve it with new algoritmhms.
\end{document}